MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-11 chapters/MetaTracts.tex
dell8 local changes
Updated images

\begin{figure}[tb]
\centering
\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
\caption{Data Characteristics. (A) shows a rendering of the data. (B) shows 2D slices along the X and (C) along the Z axis. (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$. It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}A shows a volume rendering of the dataset. B shows 2D slices along the X(right) and C in Z(left) axis. Figure~\ref{fig:data-char}D shows the zoomed in version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}D contains multiple bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}A shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

@@

%
\begin{figure}[tb]
\centering
\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
\label{fig:reliable_hessian}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in green. Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the blue arrow (Figure~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the blue dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
  \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
  \end{center}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
\end{figure}
The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
	\label{fig:length_distribution}
\end{figure}

@@

\subsection {Orientation based clustering}
%We use the inherent features expressed as weaving pattern of our data to pre-cluster the metaTracts into the number of classes based on the major directions of the pattern.
%It is important to note that this step broadly divides the MetaTracts into classes based on the major orientations and not individual fiber bundles. 
This step divides the individual MetaTracts into classes based on their major orientations. In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01} and later used in DTI fiber coloring~\cite{Brun2003} and fiber clustering~\cite{Brun2004}. The key notion is to find a suitable similarity measure to define the weights of an undirected weighted graph. The nodes being the data points and the edges being the weights. Then, an eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
Let G be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:
\begin{equation}\label{equn:eigenMaps}
L\textbf{f}=\lambda D\textbf{f}
\end{equation}
where D is the diagonal weight matrix and L is the Laplacian matrix. The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an m-dimensional space. K-means clustering is then used to cluster in this lower dimensional space.

@@

\begin{figure}[tb]
  \centering
  	%\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}~
  	%	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}
%  	\captionsetup[subfigure]{labelformat=empty}
%  	\subfloat[1]{\includegraphics[width=0.2\linewidth]{imagesMT2014/orientation_clustering_A.png}}
%  	\subfloat[2]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_1.PNG}}
% 	\subfloat[3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_2.PNG}}
% 	\subfloat[4]{\includegraphics[width=0.2\linewidth]{imagesMT2014/curveD.PNG}}
%  	
 \includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
  	\caption{(C) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts projected on top three eigenvectors. (B) 2D slices of (C). }
  \label{fig:orientation_clustering}
  \end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.

@@

\section {Experimental Results}
\label {sec:results}
%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}
\begin{figure*}
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B). (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure*}

\begin{figure}
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-6.PNG}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in red, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}

@@

 %glass
We tested our method on multiple datasets with varying characteristics. 
Figure~\ref{fig:crop-16-decomp} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively). The hierarchical clustering is robust to over segmentation.
Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster. We generate the correct clusters and outlier bundles with very few elements which can be discarded. As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length). Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle. 
Fig.~\ref{fig:prepreg} shows the results on a dense data set with \textit{flat and thin} bundles. The bundles are indistinguishable in the original data set (Fig.~\ref{fig:prepreg}A). A red dotted line shows one of the fiber bundle boundary. Fig.~\ref{fig:prepreg}B shows the fiber bundles for the two orientation. Fig.~\ref{fig:prepreg}C shows the combined results.Fig.~\ref{fig:prepreg}D shows the volume after voxelization.
The fiber bundle extraction using MetaTracts was implemented in C++ using ITK. All clustering was done in R. The preprocessing consists of MetaTract generation the distance computation and the orientation clustering. The hierarchical clustering separately, on the results of the orientation clusters can be done quickly in a few minutes on a Intel Xeon E5-2667 workstation according to user input `h'(section~\ref{sec:param_choices}). We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.

\section {User Evaluation}
\label {sec:evaluation}
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
(b) shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
\label{fig:context}
\end{figure}

>>>>

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
\includegraphics[width=0.5\textwidth]{imagesMT2014/image1}
\caption{Data Characteristics. (A) shows a rendering of the data. (B) shows 2D slices along the X and (C) along the Z axis. (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

Figure \ref{fig:data-char} shows dataset1 with woven fiber bundles. Dataset1 is 450x300x500 voxels in size with isotropic resolution of $2\mu m$, the datatype is uint8. It clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}A shows a volume rendering of the dataset. B shows 2D slices along the X(right) and C in Z(left) axis. Figure~\ref{fig:data-char}D shows the zoomed in version of the green square. 
The resolution of dataset1 renders the individual carbon fibers in a fiber bundle are hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}D contains multiple bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles themselves may differ in terms of the amount of fibers in the bundle, figure~\ref{fig:data-char}A shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. In all our data sets the number of separate orientations was two. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence of the weaving pattern of the fiber bundles, individual fibers may be adjacent in euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item (Local orientation) Each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item (Connectivity) Moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\begin{figure}[tb]
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\includegraphics[width=0.5\textwidth]{imagesMT2014/image3.png}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_rh.png}
\includegraphics[width=0.5\textwidth]{imagesMT2014/image2}
\caption{Reliable Hessians. (A) Colored according to Orientation vector mapped to RGB. (B) 2D slice along Z and (C) along X axis. (D) close up of the region marked in green in C.}
\label{fig:reliable_hessian}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in green. Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the blue arrow (Figure~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the blue dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
   \includegraphics[ width=0.45\textwidth]{imagesMT2014/image4}
 % \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
  \end{center}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
\end{figure}
The set of vertices which are in the blue region but not overlapped by green are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
		\includegraphics[width=0.45\textwidth]{imagesMT2014/image5}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
	\label{fig:length_distribution}
\end{figure}

@@

\subsection {Orientation based clustering}
%We use the inherent features expressed as weaving pattern of our data to pre-cluster the metaTracts into the number of classes based on the major directions of the pattern.
%It is important to note that this step broadly divides the MetaTracts into classes based on the major orientations and not individual fiber bundles. 
This step divides the individual MetaTracts into classes based on their major orientations. In order to cluster MetaTracts going in same directions, we use a spectral embedding technique called Laplacian eigenmaps which was originally introduced by Belkin and Niyogi \cite{Belkin01} and later used in DTI fiber coloring~\cite{Brun2003} and fiber clustering~\cite{Brun2004}. The key notion is to find a suitable similarity measure to define the weights of an undirected weighted graph. The nodes being the data points and the edges being the weights. Then, an eigenvalue problem is solved, which maps the manifold embedded in the graph into a lower dimensional space while preserving the graph structure.
This lower dimensional space can then be clustered more efficiently than the higher dimensional space. 
Let G be the graph, we compute the eigenvalues and eigenvectors for the generalized eigenvector problem:$L\textbf{f}=\lambda D\textbf{f}$,
%\begin{equation}\label{equn:eigenMaps}
%L\textbf{f}=\lambda D\textbf{f}
%\end{equation}
where D is the diagonal weight matrix and L is the Laplacian matrix. The eigenvector \textbf{${f}_{0}$} corresponding to the eigenvalue 0 is left out and the next $m$, {\textbf{${f}_{1}$} through \textbf{${f}_{m}$}} eigenvectors are used to embed in an m-dimensional space. K-means clustering is then used to cluster in this lower dimensional space.

@@

\begin{figure}[tb]
  \centering  	
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
  \includegraphics[width=0.45\textwidth]{imagesMT2014/image6}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (B) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters are shown in Fig.~\ref{fig:crop-16-decomp}.
\begin{figure*}
\centering
		\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
		\includegraphics[width=0.4\textwidth]{imagesMT2014/image7b}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of (B). (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure*}

@@

\section {Experimental Results}
\label {sec:results}
%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}

\begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8.PNG}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}

@@

 %glass
We tested our method on multiple datasets with varying characteristics. 
Figure~\ref{fig:crop-16-decomp} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively). The hierarchical clustering is robust to over segmentation.
Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster. We generate the correct clusters and outlier bundles with very few elements which can be discarded. As noted in section~\ref{subsec:dist_clustering} the tracts do not run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length). Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle. 
Fig.~\ref{fig:prepreg} shows the results on our dataset2. This dataset is dense with \textit{flat and thin} bundles. The dimensions are 300x350x300 and the datatype is uint16. The bundles are indistinguishable in the original data set Fig.~\ref{fig:prepreg}(A). A green dotted line shows one of the fiber bundle boundary. Fig.~\ref{fig:prepreg}(B) shows the fiber bundles for the two orientation. Fig.~\ref{fig:prepreg}(C) shows the combined results.Fig.~\ref{fig:prepreg}(D) shows the volume after voxelization.
The fiber bundle extraction using MetaTracts was implemented in C++ using ITK. All clustering was done in R. The preprocessing consists of MetaTract generation the distance computation and the orientation clustering. The hierarchical clustering separately, on the results of the orientation clusters can be done quickly in a few minutes on a Intel Xeon E5-2667 workstation according to user input `h'(section~\ref{sec:param_choices}). We store the normalized direction at each fiber point, thus both distance measures can be extracted simultaneously without extra computation.

\section {User Evaluation}
\label {sec:evaluation}
%\begin{figure}[h]
%\centering
%\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
%\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
%\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
%(b) shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
%\label{fig:context}
%\end{figure}
